\documentclass[t]{beamer}

\subtitle{Diophantine Equations}

\input{../../_tools/setup}

\begin{document} 
	\startdoc
	
	\topics{
		% Last Time
		\item More with gcds
		\item If $a^n|b^n$ then $a|b$
	}
	{
		% This time
		\item Diophantine Equations of the form $ax+by=c$
		\item Solving Diophantine Equations
	}

\slide{
	\begin{defn}
		The \emph{Euclidean algorithm} is a method from computing $\gcd(a,b)$. Assume for simplicity that $a\geq b>0$ since $\gcd(|a|,|b|)=\gcd(a,b)$ and $\gcd(a,0)=a$ if $a>0$. By the division algorithm, we can write
		
			\[
			\begin{array}{rcll}
			a&=&q_0b+r_1 & 0<r_1<b\\
			b&=&q_1r_1+r_2 & 0<r_2<r_1\\
			r_1&=&q_2r_2+r_3 & 0<r_3<r_2\\
			r_2&=&q_3r_3+r_4 & 0<r_4<r_3\\
			&\vdots\\
			r_{n-2}&=&q_{n-1}r_{n-1}+r_n & 0<r_n<r_{n-1}\\
			r_{n-1}&q_nr_n+0.
			\end{array}
			\]
		Then \fbox{$r_n=\gcd(a,b)$}.
	\end{defn}
}
\slide{
	\begin{statementblock}{Theorem 2.9}
		The linear Diophantine equation  $ax+by=c$ has a solution if and only if $d|c$, where $d=\gcd(a,b)$. If $x_0,y_0$ is any particular solution of this equation ,then all other solutions are given by
			\[x=x_0+\left(\frac{b}{d}\right)t,\quad y=y_0-\left(\frac{a}{d}\right)t,\]
		where $t$ is an arbitrary integer.
	\end{statementblock}
	\begin{defn}
		To \emph{solve} a Diophantine equation is to find all possible integer$^*$ values that satisfy the equation.
	\end{defn}
}
\slide{
	\begin{exercise}
		Which of the following Diophantine equations can be solved?
		\enumalph{
			\item $900x+6y=36$
			\item $300x+225y=4$
			\item $100x+90y=30$
			\item $36x+18y=50$
	}
	\end{exercise}
}
\slide{
	\begin{exercise}
		Solve the Diophantine equation
		\[100x+90y=30.\]
	\end{exercise}
}
\slide{
	\begin{exercise}
		Solve the Diophantine equation
		\[12012x+3575y=1859.\]
		
		\fbox{
		$\begin{array}{rcll}
		12012&=&3\cdot 3575+1287\\
		3575&=&2\cdot 1287+1001\\
		1287&=&1\cdot 1001 + 286\\
		1001&=&3\cdot 286+143\\
		286&=&2\cdot 143
		\end{array}
		$
		}
	\end{exercise}
}
\slide{
\begin{exercise}
	Solve this problem posed by B\'ezout in 1766.
	
	You have a bag of silver coins, each coin weighing 17 ounces of silver. A merchant has a bag of silver coins, each weighing 11 ounces. You wish to pay the merchant 542 ounces of silver.  How many 17-ounce coins should you give the merchant, and how many 11-ounce coins should the merchant give you?
	
	What is the smallest number of coins that can be exchanged in this transaction?
\end{exercise}
}
\slide{
	\begin{exercise}
		A farmer purchased 100 head of livestock for a total cost of \$4000. Prices were as follows: calves, \$120 each; lambs \$50 each; piglets \$25 each. If the farmer obtained at least one animal of each type, how many of each did they buy? 
	\end{exercise}
}
\slide{
	\begin{exercise}
		Solve this problem posed by Euler in 1770.
		
		Divide 100 into two summands such that one is divisible by 7 and the other by 11.
	\end{exercise}
}
\end{document}

