\documentclass[t]{beamer}

\subtitle{Solving Linear Congruence and the Chinese Remainder Theorem}

\input{../../_tools/setup}

\begin{document} 
	\startdoc
	
	\topics{
		% Last Time
		\item Zero Divisors
		\item Representations of Integers beyond decimal
		\item Fast exponentiation algorithm
		\item Sums of squares
		\item Sums of cubes
	}
	{
		% This time
		\item Solving Linear Congruence Equations
		\item The Chinese Remainder Theorem
	}



\slide{
	\begin{defn}
		A \emph{linear congruence} is an equation of the form
			\[ax\equiv b\pmod n,\]
		where $a,b,n\in\Z$ and $x$ is a variable.  
		A \emph{solution} to this linear congruence is an integer $x_0$ such that
			\[ax_0\equiv b\pmod n.\]
	\end{defn}
	\begin{exa} Find a solution.
		\enumalph{
			\item $3x\equiv 7\pmod{10}$\vskip .15in
			\item $2x\equiv 6\pmod{10}$\vskip .15in
			\item $2x\equiv 7\pmod{10}$
		}
	\end{exa}
}
\slide{
	\begin{defn}
		To \emph{solve} the linear congruence $ax\equiv b\pmod n$, we find all $0\leq x\leq n-1$ that satisfy the equation.
	\end{defn}
	\begin{exa} Solve.
		\enumalph{
			\item $3x\equiv 7\pmod{10}$\vskip .25in
			\item $2x\equiv 6\pmod{10}$\vskip .25in
			\item $2x\equiv 7\pmod{10}$
		}
	\end{exa}
}
\slide{
	\begin{nb}
		By definition of congruence, saying $ax\equiv b\pmod n$ is equivalent to saying there is some $y$ such that $ax-b=ny$.
	\end{nb}
}
\slide{
	\begin{statementblock}{Theorem 4.7}
		The linear congruence $ax\equiv b\pmod n$ has a solution if and only if $d|b$, where $d=\gcd(a,n)$.
		
		In that case there are exactly $d$ mutually incongruent solutions of the form
			\[x_0, x_0+\frac{n}{d}, x_0+2\frac{n}{d}+\cdots+x_0+(d-1)\frac{n}{d}.\]
	\end{statementblock}
}
\slide{
	\begin{exercise}
		Solve the linear congruence $9x\equiv 21\pmod{30}$.
	\end{exercise}
}
\slide{
	\begin{exercise}
		Solve the linear congruence $7x\equiv 3\pmod{24}$.
	\end{exercise}
}
\slide{
	\begin{statementblock}{Corollary}
		If $\gcd(a,n)=1$, then the linear congruence $ax\equiv b\pmod n$ has a unique solution modulo $n$.
	\end{statementblock}
	\begin{defn}
		Given $\gcd(a,n)=1$, the unique solution to $ax\equiv 1\pmod n$ is called the \emph{(multiplicative) inverse of $a$ modulo $n$}.
	\end{defn}
}
\slidet{Finding the inverse of $a$ modulo $n$}{
	\exer{
		Find the inverse of $13$ modulo $231$.
	}
}
\slide{
	\begin{exercise}
		Find an integer $x$ that satisfies
		\[3x\equiv7\pmod 8\quad\text{and}\quad 4x\equiv3\pmod{11}.\]
	\end{exercise}
}
\slide{
	\begin{exercise}
		Find an integer $x$ that satisfies
		\[6x\equiv4\pmod{10},\quad 3x\equiv7\pmod 8\quad\text{and}\quad 4x\equiv3\pmod{11}.\]
	\end{exercise}
}
\slide{
	\begin{statementblock}{Chinese Remainder Theorem (CRT)}
		Let $n_1,n_2,\dots,n_r$ be positive integers such that $\gcd(n_i,n_j)=1$ for all $i\neq j$.  Then the system
		\begin{eqnarray*}
		x&\equiv& a_1\pmod{n_1}\\
		x&\equiv& a_2\pmod{n_2}\\
		&\vdots\\
		x&\equiv& a_r\pmod{n_r}\\
		\end{eqnarray*}
		has a simultaneous solution which is unique modulo $N=n_1\cdot n_2\cdots n_r$.
	\end{statementblock}
}
\slide{
	\begin{proof}[``Proof''.]
		\bulletize{
			\item For all $k$, set $N_k= \frac{N}{n_k}$.
			\item For all $k$, set $x_k$ to be the solution of $N_k x\equiv 1\pmod{n_k}$.
			\item Then a solution is
				\[\bar x=a_1N_1x_1+a_2N_2x_2+\cdots+a_rN_rx_r.\]
		}
	\end{proof}
}
\slide{
	\begin{exercise}
		Solve the system
		\begin{multline*}
		x\equiv6\pmod{11},\quad x\equiv13\pmod{16},\\ x\equiv9\pmod{21},\quad x\equiv 19\pmod{25}
		\end{multline*}
	\end{exercise}
}
\slide{
	\begin{statementblock}{Lemma}
		If $m\mid n$ and $a\equiv b\pmod n$, then $a\equiv b\pmod m$ too.
	\end{statementblock}
}
\slide{
	\begin{statementblock}{Lemma}
		If \enumalph{\item$\gcd(m_1,m_2)=1$, \item$x\equiv y\pmod {m_1}$, and \item$x\equiv y \pmod{m_2}$,} then $$x\equiv y\pmod {m_1m_2}.$$
	\end{statementblock}
}
\slide{
	\begin{exercise}
		Solve the linear congruence
			\[x\equiv 9\pmod {276}\]
		by using $276=2^2\cdot 3\cdot 23$ and the CRT.
	\end{exercise}
}
\slide{
	\begin{exercise}
		Solve the linear congruences
		\[x^2\equiv 1\pmod{16}\quad\text{and}\quad x^2\equiv1\pmod 9.\]
	\end{exercise}
}
\slide{
	\begin{nb}
		We can also solve systems if the moduli are not relatively prime.
	\end{nb}
	\begin{exercise}
		Solve 
		\begin{equation*}
			x\equiv6\pmod{12},\quad x\equiv2\pmod{8}.
		\end{equation*}
	\end{exercise}
}
\slide{
	\begin{exercise}
		Solve 
		\begin{equation*}
			x\equiv6\pmod{12},\quad x\equiv5\pmod{8}.
		\end{equation*}
	\end{exercise}
}
\slide{
	\begin{nb}
		We can also solve systems if there are coefficients.
	\end{nb}
	\begin{exercise}
		Solve 
		\begin{equation*}
			5x\equiv6\pmod{12},\quad 3x\equiv2\pmod{7}.
		\end{equation*}
	\end{exercise}
}
\slide{
	\begin{statementblock}{Theorem}
		The system of linear congruences
		\[ax+by\equiv r\pmod n\]
		\[cx+dy\equiv s\pmod n\]
		has a unique solution modulo $n$ whenever $\gcd(ad-bc,n)=1$.
		
		Specifically if $t$ is an integer such that $t(ad-bc)\equiv 1\pmod n$, then the solution satisfies
			\[x\equiv t(dr-bs)\pmod n\]
			\[y\equiv t(as-cr)\pmod n.\]
	\end{statementblock}
}
\slide{\vskip-.25in
	\begin{proof}[``Proof''.]
		\bulletize{
			\item Multiply the first equation by $d$ and the second equation by $b$ and subtract to get
				\[(ad-bc)x\equiv (rd-sb)\pmod n.\]
			\item Multiply the left and right of this equation by $t$ to get
				\[x\equiv t(rd-sb)\pmod n.\]
			\item Similarly, for $y$ multiply the first equation by $c$ and second by $a$ then subtract the first from the second to get
				\[(ad-bc)y\equiv (as-cr)\pmod n.\]
			\item Multiply the left and right of this equation by $t$ to get
				\[y\equiv t(as-cr)\pmod n.\]
		}
	\end{proof}
}
\slide{
	\begin{exercise}
		Solve the system of linear congruences.
			\begin{eqnarray*}
			x-y&\equiv& 14\pmod {15}\\
			4x+3y&\equiv&10\pmod{15}
			\end{eqnarray*}
	\end{exercise}
	}
\end{document}

