\documentclass[12pt]{article}
\usepackage{amsmath,amsthm,amssymb,hyperref}
\usepackage[margin=1in]{geometry}


\newcommand{\N}{\mathbb{N}}
\newcommand{\Q}{\mathbb{Q}}
\newcommand{\R}{\mathbb{R}}
\newcommand{\Z}{\mathbb{Z}}

\newcommand{\lcm}{\operatorname{lcm}}

\begin{document}
	\newgeometry{margin=.75in}
	\begin{center}
		{\Large \bf Math 341 Midterm Exam}
	\end{center}
	\section*{Purpose}
		\begin{itemize}
			\item Summative assessment on content from the course, namely
				\begin{itemize}
					\item learning from a specific online source,
					\item an introduction to proofs,
					\item factorials,
					\item divisibility,
					\item the fundamental theorem of arithmetic, and
					\item primes.
				\end{itemize}
		\end{itemize}
	\section*{Task}
		\begin{itemize}
			\item Complete the exercises listed on the following page(s) of this document and submit your solutions as a pdf to Canvas.
			\item Your solutions must be TeX'd.
			\item You must complete the entire assignment on your own.  You may not discuss anything about the exam with your classmates until after the due date.
			\item Allowed resources: our textbook, your notes, videos linked in Canvas.
			\item Unacceptable resources: anything you find on an internet search. Do NOT use a homework help website (e.g., Chegg). Their solutions are often wrong or use incorrect context.  I want you to practice making arguments that are yours. Take some ownership.
		\end{itemize}
	\section*{Criteria}
		\begin{itemize}
			\item All short answer response questions will be graded from 0 to 2 points according to accuracy and use of full sentences.
			\item All proof exercises will be graded out of 10 using the rubric linked on Canvas. Please take some time to read through the rubric and evaluate your proofs using it.
			\item (Counter)Examples should follow the same formality as proofs. Explain in a clear sentence or two why your (counter)example works.  All (counter)examples will be graded on a scale of 0-5:
			\begin{itemize}
				\item 5 - perfect (counter)example and clear explanation
				\item 4 - (counter)example works but is not explained
				\item 3 - the (counter)example doesn't work but is on the right track
				\item 2 - either an unrelated example or an incorrect proof is given
				\item 1 - it is unclear whether a (counter)example or proof are provided
				\item 0 - no work
			\end{itemize}
		\end{itemize}
	\newpage
	\restoregeometry
\begin{center}
	Math 341 Midterm Exam\\
	YOUR NAME HERE
\end{center} 
\begin{enumerate}
	\item (20 pts) Watch \url{https://www.youtube.com/watch?v=KwFyO-ubuSM}, ProfOmarMath's video on Pell's Equation then answer the following questions.
		\begin{enumerate}
			\item (5 pts) In at least 3 sentences, describe the method that Professor Omar used to find infinitely many solutions to $x^2-2y^2=1$.
				\vfill
			\item Consider the equation $x^2-7y^2=1$.  
				\begin{enumerate}
					\item (2 pts) Find a single integer solution to this equation with $1\leq x,y\leq 10$. (Remember to give a small explanation, verifying your solution works for this and all computation problems.)
						\vskip 1in
					\item (5 pts) Find a recurrence relation that will produce infinitely many solutions the equation.
						\vfill
						\newpage
					\item (3 pts) Using your recurrence relation, write down at least 4 more solutions to the equation.
						\vfill
					\item (5 pts) Find an ratio of integers that approximates $\sqrt{7}$ to at least 4 decimal places. 
					\vfill
				\end{enumerate}
		\end{enumerate}
	\newpage
	\item (10 pts) If you can hop 289 units left and right, and skip 323 units left and right, where can you travel (beginning at 0)?  Note: 5 of these points are designated for you showing the entire Euclidean Algorithm steps.
		\vfill
		\newpage
	\item (10 pts) Prove that if $x$ is an integer for which $x|x^2+1$, then $x=\pm 1$.
		\vfill
		\newpage
	\item (20 pts) 
		\begin{enumerate}
			\item Let $r$ and $s$ be integers. Use proof by induction to show that if $n\geq 1$ is odd, then $r+s$ divides $r^n+s^n$.
			
			Note: You must use proof by induction.  Keep in mind that you can assume the inductive hypothesis for an odd integer $k$ then consider the case $n=k+2$, this is valid induction.
				\vfill
				\newpage
			\item Use the result of part (a) to show that 
				\[1^{47}+2^{47}+3^{47}+4^{47}+5^{47}+6^{47}\]
				is divisible by 7.
				
				Note: You must use part (a).  Little credit will be given for any other solution. You can though, do this part without proving part (a).
				\vfill
		\end{enumerate} 
	\newpage
	\item \begin{enumerate}
		\item (5 pts) If $x\in\Z$ what are the possible 1's digits of $x^2$? (Find and prove.)\vfill
		\item (5 pts) If $x\in\Z$ what are the possible 1's digits of $x^3$? (Find and prove.)\vfill
		\item (5 pts) If $x\in\Z$ what are the possible remainders when dividing $x^4$ by 7? (Find and prove.)\vfill
	\end{enumerate}
	\newpage
	\item (5 pts) For what values of $1\leq n\leq 20$ does
		$$x^2+x\equiv 1\pmod n$$
		have a solution?
		
		For those that have a solution, find all $0\leq x<n$ that satisfy the equivalence.
		
		If you use Sage or another programming language to solve this exercise, you must submit your code in some readable format. And really, you should be using code for this.
	\end{enumerate}
\end{document}